---
title: "Tidy Tuesday (06/06/2023): Energy"
author: "Dave Tirtariyadi"
date: "2023-06-06"
categories: [tidy tuesday]
draft: true
execute: 
  message: false
  warning: false
  cache: true
---

The dataset used here comes from Our World in Data's Energy Data Explorer. The complete dataset can be found [here](https://github.com/owid/energy-data).

```{r}
#| label: load-pkgs-data

library(tidyverse)
library(tidymodels)
library(skimr)
library(knitr)
library(kableExtra)

owid_energy <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-06-06/owid-energy.csv')

```

As usual, let's first see what's in the dataset.Looking at the `skim` and the [data dictionary](https://github.com/rfordatascience/tidytuesday/tree/master/data/2023/2023-06-06#data-dictionary), we see that the dataset contains various energy performance indicators for a list of countries at different times. Specifically, the dataset has information on `r n_distinct(owid_energy$country)` countries from `r min(owid_energy$year)` to `r max(owid_energy$year)`.

```{r}
#| label: block3

owid_energy %>% 
  filter(country %in% c("Australia", "India", "Japan")) %>% 
  ggplot(mapping = aes(x = year, y = coal_prod_change_pct)) +
  geom_line() +
  facet_wrap(vars(country))

```

